---
title: "A Simulation Exercise"
author: "NiaS"
date: "October 27, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal
The exponential distribution can be simulated in R with `rexp(n, lambda)` where
`lambda` $\lambda$ is the rate parameter. The mean of exponential distribution is 
$1/\lambda$ and the standard deviation is also $1/\lambda$. For this simulation,
we set $\lambda=0.2$. In this simulation, we investigate the distribution of
averages of 40 numbers sampled from exponential distribution with $\lambda=0.2$.


## Generated the Simulation Data
```{r}
#The function parameter
lambda <- 0.2
s <- 40
B <- 1000

set.seed(1232434)

#construct a 1000 x 40 matrix
sim <- matrix(data=rexp(s*B, lambda), nrow=B)
means_of_row <- rowMeans(sim)
```

## Sample mean vs theoritical mean
```{r, fig.height=5, fig.width=6}
library(dplyr)
library(ggplot2)
#calculate the mean for each row of the matrix
m <- data.frame(Mean = apply(sim, 1, mean))

#calculate the mean of the simulated mean
m.df <- m %>% summarize(sim.mean = mean(Mean)) %>% unlist()

m %>% ggplot(aes(x=Mean))+
    geom_histogram(alpha=0.6,  binwidth = 0.3, fill="white", color="black", mapping = aes(y = ..density..)) +
  stat_function(fun = dnorm, size = 1.3) +
  geom_vline(xintercept = m.df, color="#5499C7", size = 1.5) + geom_density(alpha=.2, fill="#FF6666")+
  ggtitle("Distribution of simulated means")+theme_classic()

```


From above plot, we can conclude that the distribution of mean is centered around the mean of our simulated distribution, i.e. indicate by the position of blue vertical line.

## Sample Variance vs Theoritical Variance
```{r}
#Calculate the standard deviation of the sample 
sample.sd <- m %>% select(Mean) %>% unlist() %>% sd()
#Calculate the variance using the value of standard deviation
sample.var <- sample.sd^2
sample.var
#Theoritical variance 
((1/lambda) / sqrt(s)) ^2

```
We can observe their value are close, which are 0.625 and 0.6142007 respectively.

##Normality of the Distribution
CLT says the distribution of averages of normalized variables becomes that of a standard normal distribution as the sample size increases. 
```{r, figure.height=5, figure.width=6}

#plot the correspondent histogram and overaly with density function from theoritical sampling distribution
xfit <- seq(min(m$Mean), max(m$Mean), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(s)))

hist(means_of_row, breaks=50, prob=TRUE,
     main="Distribution of averages of samples,
     drawn from exponential distribution with lambda=0.2",
     xlab="Mean", col = "#CACFD2")
# Density of the averages of samples
lines(density(m$Mean), col="#3498DB", lwd=4)
# Theoretical center of distribution
abline(v=1/lambda, col="#8E44AD", lwd=4)
# Theoretical density of the averages of samples

lines(xfit, yfit, pch=20, col="#F7DC6F", lty=2, lwd=4)
legend('topright', c("simulation", "theoretical"), lty=c(1,2), col=c("#3498DB", "#F7DC6F"), lwd=1.5)
```


The distribution of sample means is centered at `r mean(means_of_row)`
and the theoretical center of the distribution is $\lambda^{-1}$ = `r 1/lambda`.
The variance of sample means is `r var(means_of_row)` where the theoretical variance
of the distribution is $\sigma^2 / n = 1/(\lambda^2 n) = 1/(0.04 \times 40)$ =
`r 1/(0.04 * 40)`.

Due to the Central Limit Theorem (CLT), the averages of samples follow normal
distribution. The figure above also shows the density computed using the histogram and the
normal density plotted with theoretical mean and variance values. 

Let's evaluate the coverage for 95% CI, i.e. $1/\lambda = \bar{X} \pm 1.96 \frac{S}{\sqrt{n}}$
```{r}
library(ggplot2)

val <- seq(4, 6, by=0.01)
CI <- sapply(val, function(l) {
    mu_hats <- rowMeans(matrix(rexp(s*B, rate=0.2),
                               B, s))
    ll <- mu_hats - qnorm(0.975) * sqrt(1/lambda**2/s)
    ul <- mu_hats + qnorm(0.975) * sqrt(1/lambda**2/s)
    mean(ll < l & ul > l)
})

qplot(val, CI,color = "#E9967A", show.legend=FALSE) + geom_hline(yintercept=0.95)+theme_classic()

```

As can be seen from the plot above, for selection of $\hat{\lambda}$ around 5,
the average of the sample mean falls within the confidence interval at least 95% of the time.
Note that the true rate, $\lambda$ is 5.

## Conclusion
These analysis show that the sampling distribution of the mean of an exponential distribution with $n = 40$ observations and $\lambda = 0.2$ is approximately $N(\frac{1}{0.2}, \frac{\frac{1}{0.2}}{\sqrt{40}})$ distributed.
